---
title: "Tarea_DBT845"
author: "Paula Valenzuela"
affiliation: Pontificia Universidad Católica de Valparaíso.
 email: <paula.valenzuela.aviles@gmail.com>
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(readr)
library(survival)
library(ggplot2)
library(survminer)
library(dplyr)
library(tidyr)
library(psych)
library(lmtest)
library(knitr)
library(car)
library(stats)
library(multcompView)
library(grid)
library(gridExtra)
library(cowplot)
library(PerformanceAnalytics)
library(onewaytests)
library(nlme)
library(lsmeans)
library(coefplot)
library(corrplot)
library(lme4)
library(ggpubr)
library(rstatix)
```

# Set de datos PIE
```{r}
pheno_e <- read_excel("BD_esmoltificación.xlsx", sheet = 1)
```

# Transformar a factor o numeric

```{r}
pheno_e$Animal<-as.factor(pheno_e$Animal)
pheno_e$Center<-as.factor(pheno_e$Center)
pheno_e$Treatment<-as.factor(pheno_e$Treatment)
pheno_e$Program<-as.factor(pheno_e$Program)
pheno_e$Sampling<-as.factor(pheno_e$Sampling)
pheno_e$Sex<-as.factor(pheno_e$Sex)
pheno_e$Tank<-as.factor(pheno_e$Tank)
pheno_e$Experiment<- as.factor(pheno_e$Experiment)
pheno_e$Species<- as.factor(pheno_e$Species)
pheno_e$Muscle<-as.factor(pheno_e$Muscle)
pheno_e$Spleen<-as.factor(pheno_e$Spleen)
pheno_e$Gills<-as.factor(pheno_e$Gills)
pheno_e$Serum<-as.factor(pheno_e$Serum)
pheno_e$`Head Kidney`<-as.factor(pheno_e$`Head Kidney`)
summary(pheno_e)
```
# Frecuencia 

```{r}
table(pheno_e$Center, pheno_e$Sex)
table(pheno_e$Center, pheno_e$Program)
table(pheno_e$Program, pheno_e$Tank)
table(pheno_e$Treatment, pheno_e$Sex)
```

# Histograma de Longitud y peso de los salomindos

```{r}
pheno_e %>%
  ggplot(aes(x= `Weight (g)`)) +
  geom_histogram(col="darkblue", fill= "lightblue", position = "identity", bins = 10, alpha = 0.4 )+ labs(x="Body Weight (g)", title = " Weight (g)") 

pheno_e %>%
  ggplot(aes(x= `Long (cm)`)) +
  geom_histogram(col="darkgreen", fill= "green", position = "identity", bins = 10, alpha = 0.4 )+ labs(x=" Body Length(cm)", title = "Length(cm) ")
```
# Boxplot

```{r, warning=FALSE}

a<- pheno_e %>%
  ggplot(aes(x=Center, y= `Weight (g)`, fill=Center)) +
  geom_boxplot() +
  labs(y="Body weight (g)", x="Center", 
       subtitle="Body weight variation by Center")


b<- pheno_e %>%
  ggplot(aes(x=Program, y=`Weight (g)`, fill=Program)) +
  geom_boxplot() +
  labs(y="Body weight (g)", x="Program", 
       subtitle="Body weight variation by Program")

c<- pheno_e %>%
  ggplot(aes(x=Tank, y=`Weight (g)`, fill=Tank)) +
  geom_boxplot() +
  labs(y="Body weight (g)", x="Tank", 
       subtitle="Body weight variation by Tank") + facet_grid(rows = vars(Program))+ facet_grid(rows = vars(Program))

a
b
c

```

# Correlacion individual

```{r, warning=FALSE}
par(mfrow = c(1, 1))

my_data_e <- pheno_e[, c(16,17)]

head(my_data_e, 6)

chart.Correlation(my_data_e, histogram=FALSE, pch=19, method = c("spearman"))

```

## **Importar y explorar datos de  RNA **

```{r}
RNA_e <- read_excel("BD_esmoltificación.xlsx", sheet = 2)
```

```{r}
RNA_e$Sample <- as.factor (RNA_e$Sample)
RNA_e$Animal<-as.factor(RNA_e$Animal)
RNA_e$Treatment<-as.factor(RNA_e$Treatment)
RNA_e$Program<-as.factor(RNA_e$Program)
RNA_e$Tissue<-as.factor(RNA_e$Tissue)
RNA_e$Time<-as.factor(RNA_e$Time)
RNA_e$`Individual RNA  (ng/µl)`<- as.numeric(RNA_e$`Individual RNA  (ng/µl)`)
RNA_e$`Individual absorbance`<- as.numeric(RNA_e$`Individual absorbance`)


colnames(RNA_e)[7] <- "Individual RNA (ng/µl)"
colnames(RNA_e)[8] <- "Individual absorbance"

summary(RNA_e)
```
# Histograma

```{r}
RNA_e %>%
  ggplot(aes(x= `Individual RNA (ng/µl)`)) +
  geom_histogram(col="darkblue", fill= "lightblue", position = "identity", bins = 10, alpha = 0.4 )+ labs(x="RNA (ug/ul)", title = "Concentration RNA") 

RNA_e %>%
  ggplot(aes(x= `Individual absorbance`)) +
  geom_histogram(col="darkgreen", fill= "green", position = "identity", bins = 10, alpha = 0.4 )+ labs(x="Absorbance (260/280)", title ="Absorbance RNA" )
```
# Bloxplot

```{r, warning=FALSE}

d<- RNA_e %>%
  ggplot(aes(x=Program, y=`Individual RNA (ng/µl)`, fill=Program)) +
  geom_boxplot() +
  labs(y="Concentración RNA (ng/ul)", x="Programa", 
       subtitle="RNA Concentration  variation by Program")


 e <- RNA_e%>% ggplot(aes(x=Program, y= `Individual absorbance`, fill=Program))+ geom_boxplot() +
  labs(y="RNA Absorbance A260/A280", x="Program", 
       subtitle="RNA Absorbance  variation by Program")

 # tejidos 
 f<- RNA_e %>%
  ggplot(aes(x=Tissue, y= `Individual RNA (ng/µl)`, fill=Tissue)) +
  geom_boxplot() +
  labs(y="Concentración RNA (ng/ul)", x="Tissue", 
       subtitle="RNA Concentration  variation by Tissue")
 
 g<- RNA_e%>% ggplot(aes(x=Tissue, y= `Individual absorbance`, fill=Tissue))+ geom_boxplot() +
  labs(y="RNA Absorbance A260/A280", x="Tissue", 
       subtitle="RNA Absorbance  variation by Tissue")
 
d
e
f
g
 

```

# Correlacion individual

```{r, warning=FALSE}
par(mfrow = c(1, 1))

my_data_es <- RNA_e[, c(7,8)]

head(my_data_es, 6)

chart.Correlation(my_data_es, histogram=FALSE, pch=19, method = c("spearman"))

```

# **Importación y exploración RNA Pool**
```{r}
RNA_pool_e <- read_excel("BD_esmoltificación.xlsx", sheet = 3)
```

# Transformación de factores
```{r}
RNA_pool_e$Program<-as.factor(RNA_pool_e$Program)
RNA_pool_e$Treatment<-as.factor(RNA_pool_e$Treatment)
RNA_pool_e$Tissue<-as.factor(RNA_pool_e$Tissue)
RNA_pool_e$Time<-as.factor(RNA_pool_e$Time)
RNA_pool_e$`Pool RNA (ng/µl)`<-as.numeric(RNA_pool_e$`Pool RNA (ng/µl)`)
RNA_pool_e$`Individual absorbance`<-as.numeric(RNA_pool_e$`Individual absorbance`)


colnames(RNA_pool_e)[7] <- "Individual absorbance"
colnames(RNA_pool_e)[10] <- "Pool absorbance"

summary(RNA_pool_e)

```
# Histograma

```{r}
RNA_pool_e %>%
  ggplot(aes(x= `Pool RNA (ng/µl)`)) +
  geom_histogram(col="darkblue", fill= "lightblue", position = "identity", bins = 10, alpha = 0.4 )+ labs(x="RNA (ug/ul)", title = "Concentration pool RNA") 

RNA_pool_e %>%
  ggplot(aes(x= `Individual absorbance`)) +
  geom_histogram(col="darkgreen", fill= "green", position = "identity", bins = 10, alpha = 0.4 )+ labs(x="Absorbance pool (260/280)", title ="Absorbance pool RNA" )
```

## boxplot antes y despues de los outliers

```{r}
# tejidos 

h <- RNA_pool_e %>%
  ggplot(aes(x=Tissue, y=`Pool RNA (ng/µl)`, fill=Time)) +
  geom_boxplot() +
  labs(y="RNA Concentration (ng/ul)", x="Tissue", 
       subtitle=" RNA concentration variation by Tissue and Time before delete outliers")

i<- RNA_pool_e %>%
  ggplot(aes(x=Tissue, y=`Individual absorbance`, fill=Time)) +
  geom_boxplot() +
  labs(y=" RNA Absorbance  A260/A280", x="Tissue", 
       subtitle="RNA Absorbance  variation by Tissue before delete outliers")

h
i
```

# Identicando outliers 

```{r}
out_pool0 <- RNA_pool_e %>% filter(`Individual absorbance` >= 2.2)
out_pool1 <-  RNA_pool_e %>% filter(`Individual absorbance` <= 1.6) 

out_pool <- rbind(out_pool0, out_pool1)

out_pool %>% kable(caption = "Outliers (A260/A280): 1.6 <= Samples >= 2.2")
```

# Control de calidad sin  outliers 

```{r}
is.na(RNA_pool_e$`Individual absorbance`) <- RNA_pool_e$`Individual absorbance` >= 2.2
is.na(RNA_pool_e$`Individual absorbance`) <- RNA_pool_e$`Individual absorbance` <= 1.6

j <- RNA_pool_e %>%
  ggplot(aes(x=Tissue, y=`Pool RNA (ng/µl)`, fill=Treatment)) +
  geom_boxplot() +
  labs(y="RNA Concentration (ng/ul)", x="Tissue", 
       subtitle=" RNA concentration variation by Tissue and Time after delete outliers")

k<- RNA_pool_e %>%
  ggplot(aes(x=Tissue, y= `Individual absorbance`, fill=Treatment)) +
  geom_boxplot() +
  labs(y=" RNA Absorbance  A260/A280", x="Tissue", 
       subtitle="RNA Absorbance  variation by Tissue after delete outliers")
j
k
```
## individual correlation esmoltifación

```{r, warning=FALSE}
par(mfrow = c(1, 1))

my_data_pool_e <- RNA_pool_e[, c(6,7)]


head(my_data_pool_e, 6)

res <- cor(my_data_pool_e, use="complete.obs", method = c("spearman"))
round(res, 2)

chart.Correlation(my_data_pool_e, histogram=FALSE, pch=19, method = c("spearman"))

```

# **Importar y explorar base de datos CT**

```{r, warning=FALSE}
CT_e <- read_excel("BD_esmoltificación.xlsx", sheet = 4)
CT_gene_e<- tidyr::gather(CT_e, "Gene", "ct", 6:20)
 
CT_gene_e$ct<- as.numeric(CT_gene_e$ct)
CT_gene_e$Time<- as.factor(CT_gene_e$Time)
head(CT_gene_e)

```
# Delete outliers smoltification

2 pools of samples were eliminated due to low quality 260/280.

```{r}
dim(CT_gene_e)
CT_gene_e <-subset(CT_gene_e, Pool!="RER15_5_6_15_16" & Pool!="REQ30_7_8_17_8" )   
dim(CT_gene_e)
```

# Exploratorio de ct gen elf 

```{r}
ct_his_elf_e <- CT_gene_e %>% filter(Gene=="elf") %>% ggplot(aes(x=ct))+  geom_histogram(color="darkblue", fill="lightblue", bins=10)
ct_his_elf_e

ct_gene_elf <- CT_gene_e %>% filter(Gene=="elf") %>% ggplot(aes(x=Gene, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  coord_flip()
ct_gene_elf

ct_tissue_elf<- CT_gene_e %>% filter(Gene=="elf") %>% ggplot(aes(x=Tissue, y= ct))+
  geom_boxplot()+scale_x_discrete(labels = c('Gill','HK','M',"S"))+
  facet_wrap(~ Gene,scales = "free")
ct_tissue_elf
  
ct_gene_tissue_elf <- CT_gene_e %>% filter(Gene=="elf") %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")
ct_gene_tissue_elf    

```

# ELIMINADO OUTLIERS elf outliers

## **El pool REB15_7_8_17_8 tiene un valor de elf imposible: dice 25.17**

```{r}
dim(CT_gene_e)
CT_gene_e <-subset(CT_gene_e, Pool!="REB15_7_8_17_8" )   
dim(CT_gene_e)
```


## Exploratory ct gene

```{r}
ct_his_e <- ggplot(CT_gene_e, aes(x= ct))+
  geom_histogram(color="darkblue", fill="lightblue", bins=15)+
  facet_wrap(~ Gene,scales = "free")
ct_his_e


ct_gene_e <- ggplot(CT_gene_e, aes(x=Gene, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  coord_flip()
ct_gene_e

ct_tissue_e<- ggplot(CT_gene_e, aes(x=Tissue, y= ct))+
  geom_boxplot()+scale_x_discrete(labels = c('Gill','HK','M',"S"))+
  facet_wrap(~ Gene,scales = "free")
ct_tissue_e

ct_gene_tissue_e <- ggplot(CT_gene_e, aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")
ct_gene_tissue_e

#Head kidney

ct_gene_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney")
ct_gene_HK_e %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - HK")

#Muscule

ct_gene_M_e <- CT_gene_e %>% filter(Tissue == "Muscule") 
ct_gene_M_e %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - Muscule")

#Gills 
ct_gene_G_e <- CT_gene_e %>% filter(Tissue == "Gills") 
ct_gene_G_e %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - Gills")

#spleen
ct_gene_S_e <- CT_gene_e %>% filter(Tissue == "Spleen") 
ct_gene_S_e %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - Spleen")

# Head kidney

ct_gene_HK_e$Gene <- as.factor(ct_gene_HK_e$Gene)
  
pdf(onefile = TRUE)
for(i in 1:length(levels(ct_gene_HK_e$Gene))){
A <-   ggplot(ct_gene_HK_e[ct_gene_HK_e$Gene == levels(ct_gene_HK_e$Gene)[i],], aes(x=Time, y= ct)) + geom_boxplot(color="darkblue", fill="lightblue") + ggtitle(paste0("Expresión de genes en Riñon (CT): ",levels(ct_gene_HK_e$Gene)[i]))
grid.arrange(A)
}
dev.off()


#Muscule

ct_gene_M_e$Gene<- as.factor(ct_gene_M_e$Gene)

pdf(onefile = TRUE)
for(i in 1:length(levels(ct_gene_M_e$Gene))){
A <-   ggplot(ct_gene_M_e[ct_gene_M_e$Gene == levels(ct_gene_M_e$Gene)[i],], aes(x=Time, y= ct)) + geom_boxplot(color="darkblue", fill="lightblue") + ggtitle(paste0("Expresión de genes en músculo (CT): ",levels(ct_gene_M_e$Gene)[i]))
grid.arrange(A)
}
dev.off()

#Gills
ct_gene_G_e$Gene<- as.factor(ct_gene_G_e$Gene)

pdf(onefile = TRUE)
for(i in 1:length(levels(ct_gene_G_e$Gene))){
A <-   ggplot(ct_gene_G_e[ct_gene_G_e$Gene == levels(ct_gene_G_e$Gene)[i],], aes(x=Time, y= ct)) + geom_boxplot(color="darkblue", fill="lightblue") + ggtitle(paste0("Expresión de genes en Branquia (CT): ",levels(ct_gene_G_e$Gene)[i]))
grid.arrange(A)
}
dev.off()

#Spleen
ct_gene_S_e$Gene<- as.factor(ct_gene_S_e$Gene)

pdf(onefile = TRUE)
for(i in 1:length(levels(ct_gene_S_e$Gene))){
A <-   ggplot(ct_gene_S_e[ct_gene_S_e$Gene == levels(ct_gene_S_e$Gene)[i],], aes(x=Time, y= ct)) + geom_boxplot(color="darkblue", fill="lightblue") + ggtitle(paste0("Expresión de genes en Bazo (CT): ",levels(ct_gene_S_e$Gene)[i]))
grid.arrange(A)
}
dev.off()
```

#Head Kidney (analisis estadistico) 

```{r}
ct_gene_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney")
ct_gene_HK_e %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - HK")

ct_gene_HK_e$Gene <- as.factor(ct_gene_HK_e$Gene)


```



## ct gene correlation

```{r, warning=TRUE}
Cor_ct_gene_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney") %>% spread(key = Gene, value = ct)
Cor_ct_gene_HK_e <- Cor_ct_gene_HK_e[, -9]
Cor_ct_gene_HK_e <- Cor_ct_gene_HK_e[, -9]

# TIFF HK correlation
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_HK_e[,6:18], histogram=FALSE,pch=18, method = c("spearman")) 
mtext("Correlacion expresión de genes Riñon (CT)", side=3, line=3)
# Close device
dev.off()
```

```{r}
Cor_ct_gene_M_e <- CT_gene_e %>% filter(Tissue == "Muscule") %>% spread(key = Gene, value = ct)
Cor_ct_gene_M_e <- Cor_ct_gene_M_e[, -10]


# TIFF device
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_M_e[, 6:19], histogram=FALSE, pch=30, method = c("spearman"))
mtext("Correlacion expresión de genes Músculo (CT)", side=3, line=3)
# Close device
dev.off()
```

```{r}
Cor_ct_gene_G_e <- CT_gene_e %>% filter(Tissue == "Gills") %>% spread(key = Gene, value = ct)
Cor_ct_gene_G_e <- Cor_ct_gene_G_e[, -10]


# tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_G_e[, 6:14], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Branquia (CT)", side=3, line=3)
# Close device
dev.off()

```


```{r}
Cor_ct_gene_S_e <- CT_gene_e %>% filter(Tissue == "Spleen") %>% spread(key = Gene, value = ct)
Cor_ct_gene_S_e <- Cor_ct_gene_S_e[, -9]


# TIFF device
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_S_e[, 6:19], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Bazo (CT)", side=3, line=3)
# Close device
dev.off()
```

```{r, warning=FALSE}
Cor_ct_gene_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney") %>% spread(key = Gene, value = ct)
Cor_ct_gene_HK_e <- Cor_ct_gene_HK_e[, -9]
Cor_ct_gene_HK_e <- Cor_ct_gene_HK_e[, -9]

# TIFF HK correlation
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_HK_e[,6:18], histogram=FALSE,pch=18, method = c("spearman")) 
mtext("Correlacion expresión de genes Riñon (CT)", side=3, line=3)
# Close device
dev.off()


Cor_ct_gene_M_e <- CT_gene_e %>% filter(Tissue == "Muscule") %>% spread(key = Gene, value = ct)
Cor_ct_gene_M_e <- Cor_ct_gene_M_e[, -10]


# TIFF device
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_M_e[, 6:19], histogram=FALSE, pch=30, method = c("spearman"))
mtext("Correlacion expresión de genes Músculo (CT)", side=3, line=3)
# Close device
dev.off()

Cor_ct_gene_G_e <- CT_gene_e %>% filter(Tissue == "Gills") %>% spread(key = Gene, value = ct)
Cor_ct_gene_G_e <- Cor_ct_gene_G_e[, -10]


tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_G_e[, 6:14], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Branquia (CT)", side=3, line=3)
# Close device
dev.off()

Cor_ct_gene_S_e <- CT_gene_e %>% filter(Tissue == "Spleen") %>% spread(key = Gene, value = ct)
Cor_ct_gene_S_e <- Cor_ct_gene_S_e[, -9]


# TIFF device
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_S_e[, 6:19], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Bazo (CT)", side=3, line=3)
# Close device
dev.off()


```
## Exploratory ct gene REDUCED 

```{r correlatio_plot_reduced}
ct_gene_HK_reduced <- ct_gene_HK_e[ct_gene_HK_e$Gene %in% c("Pool","Program","Treatment","Tissue","Time","Anexina 1", "mhcii","cd4","eomes","tnf","ifn","il10","il12","il1b","inos", "mhci","cd8", "TGFb","tnf" ), ] 

ct_gene_HK_reduced$Gene <- as.factor(ct_gene_HK_reduced$Gene)
 
tiff("Gene expression HK.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')

ct_gene_HK_reduced %>% ggplot(aes(x=Time, y= ct))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="CT by gene - HK")
dev.off()
```


```{r correlatio_plot_reduced}
tiff("Correlation plot.tiff", units="in", width=5, height=4, res=300, compression = 'lzw')
chart.Correlation(Cor_ct_gene_HK_e[, c( "mhcii","cd4","eomes","tnf","ifn","il10","il12","il1b","inos", "mhci","cd8", "TGFb","tnf")], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Riñon reducido (CT)", side=3, line=3)
# Close device
dev.off()
```


```{r correlatio_plot_reduced}
HK_cor <-cor(Cor_ct_gene_HK_e[,c("anexina 1", "mhcii","cd4","eomes","tnf","ifn","il10","il12","il1b","inos", "mhci","cd8", "TGFb","tnf")])




cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

#matrix of the p-value of the correlation
#p.mat <- cor.mtest(mat)
#head(p.mat[, 1:5])


png("Correlation_plot_HK.png", width = 5, height = 4, units = 'in', res = 300)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(HK_cor, method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=0, #Text label color and rotation
        diag=FALSE 
         )
dev.off()


chart.Correlation(Cor_ct_gene_HK_e[, c("anexina 1", "mhcii","cd4","eomes","tnf","ifn","il10","il12","il1b","inos", "mhci","cd8", "TGFb","tnf")], histogram=FALSE, pch=19, method = c("spearman"))
mtext("Correlacion expresión de genes Riñon reducido (CT)", side=3, line=3)


```


```{r}
stat.test <- ct_gene_HK_reduced %>%
  group_by(Gene) %>%
  wilcox_test(ct ~ Time, ref.group = "1") %>% add_significance()
stat.test

# Box plots with p-values
stat.test <- stat.test %>% add_xy_position(x="Time")
tiff("Gene_expression_HK.tiff", units="in", width=8, height=5, res=300, compression = 'lzw')
ggboxplot(ct_gene_HK_reduced, x = "Time", y = "ct", fill = "steelblue", facet.by = "Gene", scales = "free", add = "jitter", add.params = list(size = 0.9, jitter = 0.2)) +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.05, hide.ns=T, color = "red")+
  scale_y_continuous(expand = expansion(mult = c(0.20, 0.15)))
dev.off()

png("Gene_expression_HK.png", width = 8, height = 4, units = 'in', res = 300)
ggboxplot(ct_gene_HK_reduced, x = "Time", y = "ct", fill = "steelblue", facet.by = "Gene", scales = "free", add = "jitter", add.params = list(size = 0.8, jitter = 0.1)) +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.05, hide.ns=T, color = "red")+
  scale_y_continuous(expand = expansion(mult = c(0.20, 0.15)))
dev.off()


```


## PCA ct gene
```{r, warning=FALSE}
PCA_ct_gene_HK <- Cor_ct_gene_HK_e
head(PCA_ct_gene_HK)
pca <- prcomp(PCA_ct_gene_HK[, 6:18])
biplot(x = pca, scale = 0, cex = 0.7, col = c("blue4", "brown3"))

# Cálculo de la varianza explicada acumulada 
prop_varianza <- pca$sdev^2/sum(pca$sdev^2)
prop_varianza_acum <- cumsum(prop_varianza)
ggplot(data = data.frame(prop_varianza_acum, pc = factor(1:13)),
       aes(x = pc, y = prop_varianza_acum, group = 1)) +
  geom_point() +
  geom_line() +
  geom_label(aes(label = round(prop_varianza_acum,2))) +
  theme_bw() +
  labs(x = "Componentes principales", 
       y = "Prop. varianza explicada acumulada")
```
# Delta ct calibrador by Tissue
# Intento de Iterar Delta ct calibrador.
# Aun no funciona - no Run
```{r}
Calibrador_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney")  %>% group_by(Gene) %>% summarize(media_ref=mean(ct),media_target=mean(ct),Delta_calibrador=media_target-media_ref, .groups = 'drop')

Calibrador_M_e <- CT_gene_e %>% filter(Tissue == "Muscule")  %>% group_by(Gene) %>% summarize(media_ref=mean(ct),media_target=mean(ct),Delta_calibrador=media_target-media_ref, .groups = 'drop')


Calibrador_G_e <- CT_gene_e %>% filter(Tissue == "Gills")  %>% group_by(Gene) %>% summarize(media_ref=mean(ct),media_target=mean(ct),Delta_calibrador=media_target-media_ref, .groups = 'drop')

Calibrador_S_e <- CT_gene_e %>% filter(Tissue == "Spleen")  %>% group_by(Gene) %>% summarize(media_ref=mean(ct),media_target=mean(ct),Delta_calibrador=media_target-media_ref, .groups = 'drop')


Calibrador_HK_e
Calibrador_M_e
Calibrador_G_e
Calibrador_S_e

ggplot(Calibrador_HK_e, aes(x= Delta_calibrador))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta calibrador", x="Count", 
       subtitle="Delta_calibrador HK")

ggplot(Calibrador_M_e, aes(x= Delta_calibrador))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta calibrador", x="Count", 
       subtitle=" Delta_calibrador M")

ggplot(Calibrador_G_e, aes(x= Delta_calibrador))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta calibrador", x="Count", 
       subtitle=" Delta_calibrador Gills")

ggplot(Calibrador_S_e, aes(x= Delta_calibrador))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta calibrador", x="Count", 
       subtitle=" Delta_calibrador Spleen")





```
# Calculo Delta ct test
corregir valor Delta_tests porque solo sale el valor de ct, no elf

```{r}
Test_HK_e <- CT_gene_e %>% filter(Tissue == "Head kidney")  %>% mutate(Delta_test=ct)

Test_HK_e

Test_M_e <- CT_gene_e %>% filter(Tissue == "Muscle")  %>% mutate(Delta_test=ct-elf)

Test_M

Test_G <- CT_gene %>% filter(Tissue == "Gills")  %>% mutate(Delta_test=ct-elf)

Test_G

Test_S <- CT_gene %>% filter(Tissue == "Spleen")  %>% mutate(Delta_test=ct-(elf))

Test_S

ggplot(Test_HK, aes(x= Delta_test))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta test", x="Count", 
       subtitle="Delta test HK")

ggplot(Test_M, aes(x= Delta_test))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta test", x="Count", 
       subtitle="Delta test M")

ggplot(Test_G, aes(x= Delta_test))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta test", x="Count", 
       subtitle="Delta test Gills")

ggplot(Test_S, aes(x= Delta_test))+
  geom_histogram(color="darkblue", fill="lightblue", bins=10)+
  labs(y="Delta test", x="Count", 
       subtitle="Delta test Spleen")


ggplot(Test_HK, aes(x=Time, y= Delta_test))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="Delta test HK")

ggplot(Test_M, aes(x=Time, y= Delta_test))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="Delta test M")

ggplot(Test_G, aes(x=Time, y= Delta_test))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="Delta test Gills")

ggplot(Test_S, aes(x=Time, y= Delta_test))+
  geom_boxplot(color="darkblue", fill="lightblue")+
  facet_wrap(~ Gene,scales = "free")+
  labs(subtitle="Delta test Spleen")

```


```{r, warning=FALSE}

